\documentclass{article}
\usepackage{a4wide}
\usepackage{enumerate}
\usepackage{parskip}

\newcommand{\action}[1]{%
	\mbox{}\hfill \textbf{Action:~#1}%
}

\newcommand{\word}[1]{\texttt{#1}}

\makeatletter
\renewcommand{\section}{\@startsection
	{section}{1}{0pt}{1pt}{1pt}
	{\normalfont\bfseries}
}
\makeatother

\begin{document}
\pagestyle{empty}

\begin{center}\bf\large
Forth 200\emph{x} Standards Committee \#4 \\[6pt]
Technische Universit\"{a}t Wien, Vienna, Austria. \\[6pt]
25--26 September 2008 \\[12pt]
Draft Minutes
\end{center}

\section{Participants}
\label{members}

	\begin{tabular}{lll}
		Willem		Botha		& CCS, South Africa \\
		Federico de	Ceballos	& University of Cantabria, Spain \\
		M. Anton	Ertl		& Technische Universit\"{a}t Wien, Austria & (Chair) \\
		Andrew		Haley		& Red Hat, Cambridge, England \\
		Ulrich		Hoffmann	& FH Wedel, Germany	\\
		Dr. Peter	Knaggs		& University of Exeter, England \\
		Bernd		Paysan		& Germany \\
		Stephen		Pelc		& MPE, England \\
		Dr. Bill	Stoddart	& University of Teesside, England \\
		Willi		Stricker	& FH Hannover, Germany \\
		Leon		Wagner		& Forth, Inc., USA \\
	\end{tabular}

Ulrich Hoffmann and Bernd Paysan become full members of the
committee according to the membership rules.

Andrew Haley, Willi Stricker and Leon Wagner where elected
full members of the committee on there first attendance.
(8/0/0)

\section{Minutes of last meeting}
\label{minutes}
	The draft minutes of the last meeting contain a mistake:
	They claim that the \textsf{throw-iors} proposal was referred
	for further clarification, when they where accepted.  With
	this correction the minutes where accepted (10/0/1) as a true
	and accurate representation of the meeting.

\section{Review of Forth2007.2 basis document}
\label{review}
	\begin{enumerate}[i)]
	\item \LaTeX{} source file now given in page footer.

	\item Inline rational text now in sans-serif font.

	\item Renamed ``forward'' to ``foreword''.

	\item Two new sentences added to ``Process''.

	\item Membership Rules added to ``200x Membership''.

	\item Added \textsf{X:number-prefix} proposal:
		extending section 2.2.1 Numeric notation and table 3.6 Forth 200x
		Extensions. Replacing section 3.4.1.3 Text interpreter input number
		conversion, with the addition of text to 8.3.2 Text interpreter input
		number conversion.  Moving section 12.2.2 Notation to section 2.2.1
		Numeric notation.  Adding test cases to section A.3.4.1.3 Numeric
		notation.

	\item Added \textsf{X:structures} proposal:
		extended table 3.6 Forth 200x Extensions, sections 4.1.2 Ambiguous
		conditions and 10.4.1.2 Ambiguous conditions.
		Adding section 10.3.1 Data types,
		\word{BEGIN-STRUCTURE}, \word{END-STRUCTURE}, \word{+FIELD},
		\word{CFIELD:} and \word{FFIELD:} to 10.6.2 Facility extension words
		and \word{DFFIELD:}, \word{FFIELD:} and \word{SFFIELD:} to 12.6.2
		Floating extension words.

		The rational for \word{BEGIN-STRUCTURE} contains the phrase
		``school of though'' which should be corrected to ``school of
		thought''.
		\action{PK}

	\item Added \textsf{X:throw-iors} proposal:
		extended tables 3.6 Forth 200x Extensions and 9.2 \word{THROW}
		code assignments.

	\item Tidied up source code in \textsf{X:ekeys} definitions, adding
		rationale for to \word{EKEY}.

		The text relating to the relationship between \word{KEY}, \word{KEY?},
		\word{EKEY} and \word{EKEY?} still needs work, in particular the
		persistence of the value returned.
		\action{AE/SP/PK}

	\item Corrected a number of section numbering errors in appendix A
		Rationale.

	\item Added books by Elizabeth Rather and Stephen Pelc; a URL for
		the Thinking Forth entry in the appendix B Bibliography.

	\item Reformatted appendix G Change Log to allow the separate versions
		of the basis document to appear in the table of contents.
	\end{enumerate}

\section{Consideration of proposals}
\label{CfV}

\newcommand{\entry}[4]{
	\hbox{%
		\parbox[t]{8em}{\textsf{#1}}
		\parbox[t]{24em}{#2}
		\parbox[t]{4em}{#3\centering}
		\parbox[t]{6em}{\textbf{#4}\centering}
	}%
}

~\\[-\baselineskip]
\entry{\textrm{Proposal}}{Comments}{Vote}{Action}
\entry{fp-stack}{Accepted with minor modifications}{10/0/0}{AE}

\section{Discussion of \emph{draft} proposals}
\label{RfD}
~\\[-\baselineskip]
\entry{\textrm{Proposal}}{Comments}{Status}{Action}

\entry{2value}
	{Redraft with new \word{TO}}
	{Ongoing}{AE}

\entry{directories}
	{Referred for workshop discussion}
	{Ongoing}{AE}

\entry{escaped-strings}
	{Progress on to CfV}
	{Ongoing}{AE/PK}

\entry{f-value}
	{Redraft with new \word{TO}}
	{Ongoing}{AE}

\entry{synonym}
	{Minor revision to ambiguous conditions}
	{Ongoing}{SP/PK}

\entry{extended-locals}
	{Revise \{ syntax}
	{Referred}{SP}

\entry{local-buffers}
	{Withdrawn to be replaced by a new RFD}
	{Withdrawn}{}

\entry{xchars}
	{Referred for a workshop discussion to integration into a
	 proposed localisation wordset.}
	{Ongoing}{BP/AE}


%\section{Discussion of \emph{fast track} proposals}
%\label{fasttrack}
%
%\begin{tabular}{p{6.5em}p{24em}lc}
%	Proposal	& Comments				& Status & Action \\
%
%\textsf{foreword}
%	& Extend book references
%	& Accepted
%	& PK \\
%\end{tabular}

\section{Progress Reports}
\label{reports}

\begin{enumerate}[a)]
\item Development Reports
	\begin{enumerate}[i)]
	\item \word{0} for \word{NIL}
		No progress to report.
		\action{FC/AE}

	\item 16-/32-bit signed/unsigned \word{!} and \word{@} for integers,
		bytes, octets. \\
		Some progress, but would like assistance before publication.
		\action{FC}

	\item \textsf{directory}: directory accessing in general \\
		Referred to a workshop regarding use of text substitution.
		\action{AE}
	\end{enumerate}

\item Integration of RFI0007 (\word{IMMEDIATE} vs ``compilation semantics'') \\
	No progress to report.
	\action{AE/PK}

\item Obsolescent Word set \\
	It was agreed that the words should be moved to a non-normative appendix
	rather than a new wordset (11/0/0).  A RFD should be developed for the
	next meeting.
	\action{PK}

\item Review/integration of RFI documents: \\
	No progress to report.
	\action{AE/PK}
\end{enumerate}

\section{Matters Arising}
\label{matters}
\begin{enumerate}[i)]
	\item \textsf{Cross Compiler:}
		Develop Elisabeth Rather's Cross Compiler proposal into an
		RfD. \\
		\action{BP/LW/SP}

	\item \textsf{Localisation:}
		Develop L10N proposals into RFD, with the possible integration of
		the existing \textsf{xchar} proposal.
		\action{PK/SP}

	\item The length of the meeting was considered.  It was agreed that the
		next voting meeting should be extended by a day to allow time for
		workshop discussions.
\end{enumerate}

\section{Date of Next Meeting}
\label{next}

	A review meeting to be held in conjunction with the Forth-Tagung
	in April/May.
	\action{UH}

	The next voting meeting will be held directly before the next
	euroForth conference to be held in England.  The location and dates
	are yet to be decided.  Bill Stoddart, Peter Knaggs and Andrew Haley
	are all to look into conference facilities and report back to the
	email list.
	\action{PK/BS/AH}

\end{document}
