\chapter{The optional String word set} % 17
\wordlist{string}

\begin{info}
\section{The optional String word set}
\end{info}

\section{Introduction} % 17.1

\section{Additional terms and notation} % 17.2

None.

\section{Additional usage requirements} % 17.3

Append table \ref{string:env} to table \ref{table:env}.

See: \xref{usage:env}{3.2.6 Environmental queries}.

\begin{table}[ht]
  \begin{center}
	\caption{Environmental Query Strings}
	\label{string:env}
	\begin{tabular}{p{9em}rcp{0.42\textwidth}}
		\hline\hline
		\multicolumn{2}{l}{String \hfill Value data type} & Constant? & Meaning \\
		\hline
		\texttt{STRING}		& \emph{flag}	& no	&
			string word set present \\
		\texttt{STRING-EXT}	& \emph{flag}	& no	&
			string extensions word set present \\
		\hline\hline
	\end{tabular}
  \end{center}
\end{table}

\section{Additional documentation requirements} % 17.4

None.

\section{Compliance and labeling} % 17.5

\subsection{Compliance and labeling} % 17.5.1

The phrase ``Providing the String word set'' shall be appended to
the label of any Standard System that provides all of the String
word set.

The phrase ``Providing \emph{name(s)} from the String Extensions
word set'' shall be appended to the label of any Standard System
that provides portions of the String Extensions word set.

The phrase ``Providing the String Extensions word set'' shall be
appended to the label of any Standard System that provides all of
the String and String Extensions word sets.

\subsection{ANS Forth programs} % 17.5.2

The phrase ``Requiring the String word set'' shall be appended to
the label of Standard Programs that require the system to provide
the String word set.

The phrase ``Requiring \emph{name(s)} from the String Extensions
word set'' shall be appended to the label of Standard Programs that
require the system to provide portions of the String Extensions
word set.

The phrase ``Requiring the String Extensions word set'' shall be
appended to the label of Standard Programs that require the system
to provide all of the String and String Extensions word sets.


\section{Glossary} % 17.6

\begin{info}
\subsection{Glossary}
\end{info}

\subsection{String words} % 17.6.1


\begin{newword}{0170}{-TRAILING}[dash-trailing]
	\stack{c-addr $u_1$}{c-addr $u_2$}

	If $u_1$ is greater than zero, $u_2$ is equal to $u_1$ less the
	number of spaces at the end of the character string specified by
	\emph{c-addr} $u_1$. If $u_1$ is zero or the entire string
	consists of spaces, $u_2$ is zero.
\end{newword}


\begin{newword}{0245}{/STRING}[slash-string]
	\stack{c-addr$_1$ $u_1$ n}{c-addr$_2$ $u_2$}

	Adjust the character string at \emph{c-addr}$_1$ by $n$
	characters. The resulting character string, specified by
	\emph{c-addr}$_2$ $u_2$, begins at \emph{c-addr}$_1$ plus $n$
	characters and is $u_1$ minus $n$ characters long.

	\begin{rationale}
		\word{/STRING} is used to remove or add characters relative
		to the ``left'' end of the character string. Positive values
		of $n$ will exclude characters from the string while negative
		values of $n$ will include characters to the left of the
		string. \word{/STRING} is a natural factor of \word[core]{WORD}
		and commonly available.
	\end{rationale}
\end{newword}


\begin{newword}{0780}{BLANK}
	\stack{c-addr u}{}

	If $u$ is greater than zero, store the character value for
	space in $u$ consecutive character positions beginning at
	\emph{c-addr}.
\end{newword}


\begin{newword}{0910}{CMOVE}[c-move]
	\stack{c-addr$_1$ c-addr$_2$ u}{}

	If $u$ is greater than zero, copy $u$ consecutive characters
	from the data space starting at \emph{c-addr}$_1$ to that
	starting at \emph{c-addr}$_2$, proceeding character-by-character
	from lower addresses to higher addresses.

\item[Contrast with:]
	\wref{string:CMOVEtop}{CMOVE>}.

	\begin{rationale}
		If \emph{c-addr}$_2$ lies within the source region (i.e., when
		\emph{c-addr}$_2$ is not less than \emph{c-addr}$_1$ and
		\emph{c-addr}$_2$ is less than the quantity \emph{c-addr}$_1$
		$u$ \word[core]{CHARS} \word[core]{+}), memory propagation
		occurs.

		Typical use: Assume a character string at address
		100: ``ABCD''. Then after 
		\begin{quote}\ttfamily
			100 \word[core]{DUP} ~ \word[core]{CHAR+} ~ 3 \word{CMOVE}
		\end{quote}
		the string at address 100 is ``AAAA''.

		Rationale for \word{CMOVE} and \word{CMOVEtop} follows
		\word[core]{MOVE}.
	\end{rationale}
\end{newword}


\begin{newword}[CMOVEtop]{0920}{CMOVE>}[c-move-up]
	\stack{c-addr$_1$ c-addr$_2$ u}{}

	If $u$ is greater than zero, copy $u$ consecutive characters
	from the data space starting at \emph{c-addr}$_1$ to that
	starting at \emph{c-addr}$_2$, proceeding character-by-character
	from higher addresses to lower addresses.

\item[Contrast with:]
	\wref{string:CMOVE}{CMOVE}.

	\begin{rationale}
		If \emph{c-addr}$_1$ lies within the destination region (i.e.,
		when \emph{c-addr}$_1$ is greater than or equal to
		\emph{c-addr}$_2$ and \emph{c-addr}$_2$ is less than the
		quantity \emph{c-addr}$_1$ $u$ \word[core]{CHARS}
		\word[core]{+}), memory propagation	occurs.

		Typical use: Assume a character string at address
		100: ``ABCD''. Then after
		\begin{quote}\ttfamily
			100 \word[core]{DUP} \word[core]{CHAR+} \word[core]{SWAP}
			3 \word{CMOVEtop}
		\end{quote}
		the string at address 100 is ``DDDD''.
	\end{rationale}
\end{newword}


\begin{newword}{0935}{COMPARE}
	\stack{c-addr$_1$ $u_1$ c-addr$_2$ $u_2$}{n}

	Compare the string specified by \emph{c-addr}$_1$ $u_1$ to the
	string specified by \emph{c-addr}$_2$ $u_2$. The strings are
	compared, beginning at the given addresses, character by
	character, up to the length of the shorter string or until a
	difference is found. If the two strings are identical, $n$ is
	zero. If the two strings are identical up to the length of the
	shorter string, $n$ is minus-one (-1) if $u_1$ is less than $u_2$
	and one (1) otherwise. If the two strings are not identical up to
	the length of the shorter string, $n$ is minus-one (-1) if the
	first non-matching character in the string specified by
	\emph{c-addr}$_1$ $u_1$ has a lesser numeric value than the
	corresponding character in the string specified by
	\emph{c-addr}$_2$ $u_2$ and one (1) otherwise.

	\begin{rationale}
		Existing Forth systems perform string comparison operations
		using words that differ in spelling, input and output
		arguments, and case sensitivity. One in widespread use
		was chosen.
	\end{rationale}
\end{newword}


\begin{newword}{2191}{SEARCH}
	\stack{c-addr$_1$ $u_1$ c-addr$_2$ $u_2$}{c-addr$_3$ $u_3$ flag}

	Search the string specified by \emph{c-addr}$_1$ $u_1$ for the
	string specified by \emph{c-addr}$_2$ $u_2$. If \emph{flag} is
	true, a match was found at \emph{c-addr}$_3$ with $u_3$
	characters remaining. If \emph{flag} is false there was no match
	and \emph{c-addr}$_3$ is \emph{c-addr}$_1$ and $u_3$ is $u_1$.

	\begin{rationale}
		Existing Forth systems perform string searching operations
		using words that differ in spelling, input and output
		arguments, and case sensitivity. One in widespread use was
		chosen.
	\end{rationale}
\end{newword}


\begin{newword*}{2212}{SLITERAL}
\item[Interpretation:]
	Interpretation semantics for this word are undefined.

\item[Compilation:]
	\stack{c-addr$_1$ u}{}

	Append the run-time semantics given below to the current
	definition.

\item[Run-time:]
	\stack{}{c-addr$_2$ u}

	Return \emph{c-addr}$_2$ $u$ describing a string consisting of
	the characters specified by \emph{c-addr}$_1$ $u$ during
	compilation. A program shall not alter the returned string.

	\begin{rationale}
		The current functionality of \wref{core:Sq}{S"} may be
		provided by the following definition:
		\setwordlist{core}
		\begin{quote}\ttfamily
			\word{:} \word{Sq} \word{p} "ccc<quote>" -- ) \\
			\tab \word{[CHAR]} " \word{PARSE} ~
				\word{POSTPONE} \word[string]{SLITERAL} \\
			\word{;} \word{IMMEDIATE}
		\end{quote}
		\setwordlist{string}
	\end{rationale}

\end{newword*}


\subsection{String extension words} % 17.6.2

None.
